---
title: "Meeting 8: Communicate"
date: "October 28, 2021 from 7:00-8:30pm ET"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Chapters to Read

This week, we'll be discussing part five of the book, Communicate.  

- R Markdown
- Graphics for communication
- R Markdown formats
- R Markdown workflow

### Physical Book

If you're reading the physical book, the chapters to read are: 

```{r echo=FALSE}
library(tidyverse)
read_csv("../physical-book-chapter-schedule.csv") %>% 
  filter(Meeting == 8) %>% 
  select(-Meeting) %>% 
  knitr::kable()
```

### Online Book

If you're reading the online book, the chapters to read are: 

```{r echo=FALSE}
library(tidyverse)
read_csv("../online-book-chapter-schedule.csv") %>% 
  filter(Meeting == 8) %>% 
  select(-Meeting) %>% 
  knitr::kable()
```

## Exercises 

All exercises refer to the online book chapters. The book exercises and online exercises may differ, so to make sure everyone is doing the same exercises, please refer to the online book chapters.

- Chapter 27: Section 27.2.1 # 2
- Chapter 27: Section 27.3.1 # 2
- Chapter 27: Section 27.4.7 # 1
- Chapter 28: Section 28.2.1 # 3 or # 2
- Chapter 28: Section 28.3.1 # 3
- Chapter 28: Section 28.4.4 # 3

Exercises are reproduced below.

### Chapter 27 Exercises

####  Section 27.2.1 # 2

2. Create a new R Markdown document with *File > New File > R Markdown…* Knit it by clicking the appropriate button. Knit it by using the appropriate keyboard short cut. Verify that you can modify the input and see the output update.

####  Section 27.3.1 # 2

2. Using the R Markdown quick reference, figure out how to:

- Add a footnote.
- Add a horizontal rule.
- Add a block quote.

####  Section 27.4.7 # 1

1. Add a section that explores how diamond sizes vary by cut, colour, and clarity. Assume you’re writing a report for someone who doesn’t know R, and instead of setting `echo = FALSE` on each chunk, set a global option.
    
### Chapter 28 Exercises

#### Section 28.2.1 # 3 or # 2

2\. The `geom_smooth()` is somewhat misleading because the hwy for large engines is skewed upwards due to the inclusion of lightweight sports cars with big engines. Use your modelling tools to fit and display a better model.

3\. Take an exploratory graphic that you’ve created in the last month, and add informative titles to make it easier for others to understand.

#### Section 28.3.1 # 3

3. How do labels with `geom_text()` interact with faceting? How can you add a label to a single facet? How can you put a different label in each facet? (Hint: think about the underlying data.)

#### Section 28.4.4 # 3

3. Change the display of the presidential terms by:

- Combining the two variants shown above.
- Improving the display of the y axis.
- Labelling each term with the name of the president.
- Adding informative plot labels.
- Placing breaks every 4 years (this is trickier than it seems!).
