---
title: "Meeting 4: Data Importing, Tidying, and Joining"
date: "August 26, 2021 from 7:00-8:30pm ET"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Chapters to Read

This week, we'll be discussing: 

- Data Import
- Data Tidying
- Relational data

### Physical Book

If you're reading the physical book, the chapters to read are: 

```{r echo=FALSE}
library(tidyverse)
read_csv("../physical-book-chapter-schedule.csv") %>% 
  filter(Meeting == 4) %>% 
  select(-Meeting) %>% 
  knitr::kable()
```

### Online Book

If you're reading the online book, the chapters to read are: 

```{r echo=FALSE}
library(tidyverse)
read_csv("../online-book-chapter-schedule.csv") %>% 
  filter(Meeting == 4) %>% 
  select(-Meeting) %>% 
  knitr::kable()
```

## Exercises 

All exercises refer to the online book chapters. The book exercises and online exercises may differ, so to make sure everyone is doing the same exercises, please refer to the online book chapters.

- Chapter 11: Section 11.2.2 # 1, 2
- Chapter 11: Section 11.3.5 # 2, 5, 7
- Chapter 12: Section 12.2.1 # 2 (all parts)
- Chapter 12: Section 12.3.3 # 1, 3
- Chapter 12: Section 12.4.3 # 1, 2
- Chapter 12: Section 12.5.1 # 1, 2
- Chapter 12: Section 12.6.1 # 3, 4
- Chapter 13: Section 13.2.1 # 1, 4
- Chapter 13: Section 13.3.1 # 2 (all parts)
- Chapter 13: Section 13.4.6 # 1, 2, 5
- Chapter 13: Section 13.5.1 # 1, 2, 3, 5

Exercises are reproduced below.

### Chapter 11 Exercises 

```{r}
library(readr)
```


#### Section 11.2.2 (#1, #2)

1. What function would you use to read a file where fields were separated with "|"?

> Sam's Solution: 

```{r, eval = FALSE}
read_delim(file, delim = "|")
```


2. Apart from `file`, `skip`, and `comment`, what other arguments do `read_csv()` and `read_tsv()` have in common?

> Sam's Solution: 

```{r}
args(read_csv)
args(read_tsv)
```

> Brigid's question 

11.2.2 #4 

Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By default, read_csv() assumes that the quoting character will be ". What argument to read_csv() do you need to specify to read the following text into a data frame?

```{r}
txt <- "x,y\n1,'a,b'"
read_csv(txt, quote = "\'")
```


#### Section 11.3.5 (#2, #5, #7)

2. What happens if you try and set `decimal_mark` and `grouping_mark` to the same character? What happens to the default value of `grouping_mark` when you set `decimal_mark` to `","`? What happens to the default value of `decimal_mark` when you set the `grouping_mark` to `"."`?

> Live solution: 

```{r}
library(readr)
parse_double("1,234.789", 
             locale = locale(grouping_mark = ",",
                             decimal_mark = "."))
parse_double("1,234.789")
parse_number("1,234.789", 
             locale = locale(grouping_mark = ",",
                             decimal_mark = "."))
parse_double("1234.789", 
             locale = locale(grouping_mark = ",",
                             decimal_mark = "."))
# parse_double only wants you to use one mark? 

# gwynn's solution
parse_number("123,456.789", locale = locale(decimal_mark = ","))
```


> Sam's Solution: 

```{r, error=TRUE}
parse_number("123,456,789.02", 
             locale = locale(decimal_mark = ",", 
                             grouping_mark = ","))
parse_number("123.456.789,02", 
             locale = locale(grouping_mark = "."))
parse_number("123,456,789.02", 
             locale = locale(grouping_mark = "."))
```


5. What’s the difference between `read_csv()` and `read_csv2()`?

> Sam's Solution: 

```{r, eval = FALSE}
# read the Description of the read_csv function help page
?read_csv
```


7. Generate the correct format string to parse each of the following dates and times:

```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"
```

> Live Solution: 

```{r}
parse_time(t2)

parse_time(t2, "%H:%M:%OS %p")

parse_time("11:15:10.12 AM",  "%H:%M:%OS %p")
# reference 
# ?strptime
```


> Sam's Solution: 

```{r}
# ?strptime
parse_date(d1, format = "%B %d, %Y")
parse_date(d2, format = "%Y-%b-%d")
parse_date(d3, format = "%d-%b-%Y")
parse_date(d4, format = "%B %d (%Y)")
parse_date(d5, format = "%m/%d/%y")
parse_date(d5, format = "%D")

parse_time(t1, format = "%H%M")
parse_time(t2, format = "%H:%M:%OS %p")
```


### Chapter 12 Exercises

```{r}
library(tidyr)
```


#### Section 12.2.1 (#2, all parts)

2. Compute the rate for table2, and table4a + table4b. You will need to perform four operations:

    1. Extract the number of TB cases per country per year.
    2. Extract the matching population per country per year.
    3. Divide cases by population, and multiply by 10000.
    4. Store back in the appropriate place.
  
Which representation is easiest to work with? Which is hardest? Why?

```{r}
head(table2) 
head(table4a)
head(table4b)
```

> Sam's Solution: 

Table 2: 
```{r}
table2 %>% 
  # 1. Extract the number of TB cases per country per year.
  # 2. Extract the matching population per country per year.
  pivot_wider(names_from = type, values_from = count) %>% 
  # 3. Divide cases by population, and multiply by 10000.
  # 4. Store back in the appropriate place.
  mutate(rate = 10000 * cases / population)
```

Table 4a + 4b:

```{r}
table4a %>% 
  pivot_longer(-country, names_to = "year", values_to = "cases") %>% 
  left_join(
    table4b %>% 
      pivot_longer(-country, names_to = "year", values_to = "population")
  ) %>% 
  mutate(rate = 10000* cases / population)
```


#### Section 12.3.3 (#1, #3)

1. Why are `pivot_longer()` and `pivot_wider()` not perfectly symmetrical? Carefully consider the following example:

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return")
```

(Hint: look at the variable types and think about column names.)

`pivot_longer()` has a names_ptypes argument, e.g.  `names_ptypes = list(year = double())`. What does it do?


> Sam's Solution: 

```{r}
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return", names_ptypes =numeric())
```


3. What would happen if you widen this table? Why? How could you add a new column to uniquely identify each value?

```{r}
people <- tribble(
  ~name,             ~names,  ~values,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```

> Live Solution: 

```{r}
people %>% 
  pivot_wider(names_from = names, values_from = values)

people %>% 
  group_by(name, names) %>% 
  mutate(sample_id = row_number()) %>% 
  pivot_wider(names_from = names, values_from = values)
```


> Sam's Solution: 

```{r}
# multiple values for Phillip's age 
people %>% 
  pivot_wider(names_from = names, values_from = values) 

people %>% 
  group_by(name, names) %>% 
  mutate(sample_id = row_number()) %>% 
  pivot_wider(names_from = names, values_from = values)
```


#### Section 12.4.3 (#1, #2)

1. What do the extra and fill arguments do in `separate()`? Experiment with the various options for the following two toy datasets.

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```

> Sam's Solution:

```{r, warning=TRUE}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "warn")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "drop")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "warn")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "right")
```


2. Both `unite()` and `separate()` have a remove argument. What does it do? Why would you set it to `FALSE`?

> Sam's Solution:

```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "right", remove = FALSE)
```


#### Section 12.5.1 (#1, #2) 

1. Compare and contrast the `fill` arguments to `pivot_wider()` and `complete()`.

> Sam's Solution

```{r}
# values_fill is an argument in pivot_wider() that you give a scalar value that specifies what to do when there is missing data 

# example from the complete() documentation: 
df <- tibble(
  group = c(1:2, 1),
  item_id = c(1:2, 2),
  item_name = c("a", "b", "b"),
  value1 = 1:3,
  value2 = 4:6
)
# fill takes a list 
df %>% complete(group, nesting(item_id, item_name), 
                fill = list(value1 = 0, value2 = 99))

# note: pivot_wider's fill can also take a list! but complete's can't
people %>% 
  group_by(name, names) %>% 
  mutate(sample_id = row_number()) %>% 
  pivot_wider(names_from = names, values_from = values, values_fill = list(values = 999))
```


2. What does the direction argument to `fill()` do?

> Sam's Solution

```{r}
# example from tidyr::fill help 
sales <- tibble::tribble(
  ~quarter, ~year, ~sales,
  "Q1",    2000,    66013,
  "Q2",      NA,    69182,
  "Q3",      NA,    53175,
  "Q4",      NA,    21001,
  "Q1",    2001,    46036,
  "Q2",      NA,    58842,
  "Q3",      NA,    44568,
  "Q4",      NA,    50197,
  "Q1",    2002,    39113,
  "Q2",      NA,    41668,
  "Q3",      NA,    30144,
  "Q4",      NA,    52897,
  "Q1",    2004,    32129,
  "Q2",      NA,    67686,
  "Q3",      NA,    31768,
  "Q4",      NA,    49094
)

sales %>% 
  fill(year) %>% 
  head()

sales %>% 
  fill(year, .direction = "up") %>% 
  head()

sales %>% 
  fill(year, .direction = "updown") %>% 
  head()

sales %>% 
  fill(year, .direction = "downup") %>% 
  head()
```


#### Section 12.6.1 (#3, #4)

3. I claimed that `iso2` and `iso3` were redundant with country. Confirm this claim.

```{r}
who %>% 
  count(country, iso2, iso3)

length(unique(who$country))
length(unique(who$iso2))
length(unique(who$iso3))
```


4. For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.

> Live Solution

```{r}
who %>%
  pivot_longer(
    cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  ) %>% 
  mutate(
    key = stringr::str_replace(key, "newrel", "new_rel")
  ) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1) -> who5

who5 %>% 
  count(var)

who5 %>% 
  group_by(country, year, sex) %>% 
  summarize(total_cases = sum(cases)) %>% 
  filter(country == "Afghanistan") %>% 
  ggplot(aes(x = year, y = total_cases, color = sex)) + 
  geom_line()

who5 %>% 
  group_by(country, year, sex) %>% 
  summarize(total_cases = sum(cases)) %>% 
  unite(country_sex, country, sex, remove = FALSE) %>% 
  ggplot(aes(x = year, y = total_cases, color = sex, 
             group = country_sex)) + 
  geom_line(alpha = .5)

who5 %>% 
  group_by(country, year, sex) %>% 
  summarize(total_cases = sum(cases)) %>% 
  unite(country_sex, country, sex, remove = FALSE) %>% 
  ggplot(aes(x = year, y = total_cases, color = sex, 
             group = country)) + 
  geom_line(alpha = .5)
```


> Sam's Solution: 

```{r}
# we'll work with who5 from the book
who5 <- who %>% 
  pivot_longer(
    cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  ) %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel")) %>% 
  separate(key, c("new", "type", "sexage"), sep = "_") %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)

who5 %>%
  group_by(country, year, sex) %>% 
  summarize(total_cases = sum(cases)) %>% 
  unite(country_sex, country, sex, remove = FALSE) %>% 
  ggplot(aes(x = year, y = total_cases, group = country_sex, color = sex)) + 
  geom_line()

# more complex solution 
continents <- countrycode::codelist %>% select(continent, country = country.name.en)
who5 %>%
  group_by(country, year, sex) %>% 
  summarize(total_cases = sum(cases)) %>% 
  left_join(continents) %>% 
  unite(country_sex, country, sex, remove = FALSE) %>% 
  ggplot(aes(x = year, y = total_cases, group = country_sex, color = sex)) + 
  geom_line() + 
  facet_wrap(~continent, scales = "free")

# not very good, a lot of NAs. Need to look more precisely at the names in the codelist data 
```



### Chapter 13 Exercises 

#### Section 13.2.1 (#1, #4)

1. Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

> Sam's Solution 

```{r}
# You would need the coordinates of the origin and destination airports for each flight 
# found in the airports data 
# join the flights data to airports by origin and then join that data with the airports on the destination again

```


4. We know that some days of the year are "special", and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

> Sam's Solution:

```{r}
# could have a data frame with a date column and a holiday column e.g.

# month , day , holiday
# 12, 25, christmas
# 12, 31, new year's eve

# primary keys are month, day 
# connect to the flights table on month-day 
```


#### Section 13.3.1 (#2, all parts)

2. Identify the keys in the following datasets:

    - Lahman::Batting,
    - babynames::babynames
    - nasaweather::atmos
    - fueleconomy::vehicles
    - ggplot2::diamonds
    
(You might need to install some packages and read some documentation.)

> Sam's Solution: 

```{r}
glimpse(Lahman::Batting)
Lahman::Batting %>% 
  count(playerID, yearID, stint) %>% 
  filter(n > 1)
# playerID, yearID, and stint form the key 

glimpse(babynames::babynames)
babynames::babynames %>% 
  count(year, sex, name) %>% 
  filter(n > 1)
# year, sex, name form the key 

glimpse(nasaweather::atmos)
nasaweather::atmos %>% 
  count(lat, long, year, month) %>% 
  filter(n > 1)
# lat, long, year, month form the key 

glimpse(fueleconomy::vehicles)
fueleconomy::vehicles %>% 
  count(id) %>% 
  filter(n > 1)

# id is the primary key 

glimpse(ggplot2::diamonds)
ggplot2::diamonds %>% 
  # count(x, y, z, price) %>%
  count(carat, cut, color, clarity, depth, table, price, x, y, z) %>% 
  filter(n > 1)
# there is no key! 
```


#### Section 13.4.6 (#1, #2, #5)

1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

```{r, eval = FALSE}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```


(Don’t worry if you don’t understand what `semi_join()` does — you’ll learn about it next.)

You might want to use the size or colour of the points to display the average delay for each airport.

> Live Solution: 

```{r}
data("flights", package = "nycflights13")
avg_delays <- flights %>% 
  group_by(dest) %>% 
  summarize(avg_delay = mean(arr_delay, na.rm = TRUE))

data("airports", package = "nycflights13")
avg_delays %>% 
  left_join(airports, by = c("dest" = "faa")) %>% 
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(color = avg_delay), size = 1.5) +
    coord_quickmap() + 
    scale_color_gradient2(low = "navy", high = "red")
```


> Sam's Solution: 

```{r}
library(nycflights13)
flights %>%
  group_by(dest) %>% 
  summarize(mean_arr_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  left_join(airports, c("dest" = "faa")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(color = mean_arr_delay)) +
    coord_quickmap() + 
    scale_color_gradient2(high = "red", low = "green") + 
  theme_classic() + 
  theme(legend.position = "top")
```

2. Add the location of the origin and destination (i.e. the `lat` and `lon`) to 
flights.

> Live solution 

```{r}
airport_coord <- airports %>% select(faa, lat, lon)

flights %>% 
  left_join(airport_coord , by = c("origin" = "faa")) %>% 
  left_join(airport_coord, by = c("dest" = "faa"), 
            suffix = c("_origin", "_dest"))
```


> Sam's Solution: 

```{r}
flight_locations <- flights %>% 
  left_join(airports %>% select(faa, lat, lon), c("origin" = "faa")) %>% 
  left_join(airports %>% select(faa, lat, lon), c("dest" = "faa"), suffix = c("_origin", "_dest"))
```


5. What happened on June 13, 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.

> Live solution: 

```{r}
flights %>% 
  filter(month == 6, day == 13) -> june13  

june13 %>% 
  left_join(airport_coord, by = c("dest" = "faa")) %>% 
  ggplot(aes(lon, lat, color = arr_delay)) +
    borders("state") +
    geom_point(size = 2) +
    coord_quickmap()

```


> Sam's Solution: 

```{r}
flight_locations %>% 
  filter(month == 6, day == 13) %>% 
  ggplot(aes(lon_dest, lat_dest)) +
    borders("state") +
    geom_jitter(aes(color = arr_delay)) +
    coord_quickmap() 
# looks like most of the west coast flights made it but the midwest flights didn't 
# midwest means tornado
# google "June 13, 2013 tornado"
# https://en.wikipedia.org/wiki/June_12%E2%80%9313,_2013_derecho_series
```


#### Section 13.5.1 (#1, #2, #3, #5)

1. What does it mean for a flight to have a missing `tailnum`? What do the tail numbers that don’t have a matching record in `planes` have in common? (Hint: one variable explains ~90% of the problems.)

> Sam's Solution:

```{r}
flights %>% 
  filter(is.na(tailnum)) %>% 
  head()
# tailnum missing also have arr_time and other variables missing 
# flights where tailnum is missing are probably flights that were cancelled because they couldn't get a plane. 

flights %>% 
  anti_join(planes, by = "tailnum") %>% 
  count(carrier, sort = T) %>% 
  mutate(perc = n / sum(n))
# majority are from airlines MQ or AA
```


2. Filter flights to only show flights with planes that have flown at least 100 flights.

> Sam's Solution: 

```{r}
gr_100 <- flights %>% 
  filter(!is.na(tailnum)) %>% 
  count(tailnum) %>% 
  filter(n >= 100)

flights %>% 
  semi_join(gr_100, by = "tailnum")
```


3. Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models.

> Sam's Solution:

```{r}
fueleconomy::vehicles %>% 
  semi_join(fueleconomy::common, by = c("model", "make"))
```


5. What does `anti_join(flights, airports, by = c("dest" = "faa"))` tell you? What does `anti_join(airports, flights, by = c("faa" = "dest"))` tell you?

> Sam's solution:  

```{r}
anti_join(flights, airports, by = c("dest" = "faa")) %>% glimpse()
# destination code is not in airports 

anti_join(airports, flights, by = c("faa" = "dest")) %>% glimpse()
# airports which are not destinations in the data 
```


## Open Disussion

`->` and `<-`

```{r, eval = FALSE}
# "to the variable x, assign the value 4"
x <- 4
# "set x equal to 4" 
x = 4
# "assign the value 4 to the variable x"
4 -> x
# error! = behaves as <- , not as -> 
4=x
```

The assignment arrow `->` or `<-` is an arrow that points a value to a variable and assigns the value to the variable. 


`is.na` and `na.rm`: `is.na` is a function that returns `TRUE` if the value is `NA` or `NaN`, and returns `FALSE` otherwise. `na.rm` is an argument in some functions (e.g. `mean`, `sd`, etc.) which tells the function to remove any `NA` values before performing any calculations. Otherwise, if there is an `NA`, the computation will result in `NA`.  

```{r}
# ?is.na
# mean(x, na.rm = T) 
```




