---
title: "Meeting 3: Data Transformation & Exploratory Data Analysis"
date: "August 12, 2021 from 7:00-8:30pm ET"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
```

## Chapters to Read

This week, we'll be discussing: 

- Pipes
- Data Transformation
- Exploratory Data Analysis (EDA)

### Physical Book

If you're reading the physical book, the chapters to read are: 

```{r echo=FALSE}
library(tidyverse)
read_csv("../physical-book-chapter-schedule.csv") %>% 
  filter(Meeting == 3) %>% 
  select(-Meeting) %>% 
  knitr::kable()
```

### Online Book

If you're reading the online book, the chapters to read are: 

```{r echo=FALSE}
library(tidyverse)
read_csv("../online-book-chapter-schedule.csv") %>% 
  filter(Meeting == 3) %>% 
  select(-Meeting) %>% 
  knitr::kable()
```

## Exercises 

All exercises refer to the online book chapters. The book exercises and online exercises may differ, so to make sure everyone is doing the same exercises, please refer to the online book chapters.

- Chapter 5: Section 5.2.4 # 1 (odd-numbered sub-parts only), # 2 - 4
- Chapter 5: Section 5.3.1 # 1, 3, 4
- Chapter 5: Section 5.4.1 # 1 - 4
- Chapter 5: Section 5.5.2 # 1, 3, 5
- Chapter 5: Section 5.6.7 # 2 - 6
- Chapter 5: Section 5.7.1 # 2, 4, 6, 8
- Chapter 7: Section 7.3.4 # 2 - 4
- Chapter 7: Section 7.4.1 # 1, 2
- Chapter 7: Section 7.5. Do the first 2 problems of each section 7.5.1.1, 7.5.2.1, and 7.5.3.1
- BONUS: Follow the style of [this tweet](https://twitter.com/hadleywickham/status/1359852563726819332?s=20) to "translate" some of your favorite song lyrics into {tidyverse}-style code  


Exercises are reproduced below.


