---
title: "Meeting 1: Introductions to one another and to R & RStudio"
date: "July 15, 2021 from 7:00-8:30pm ET"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Chapters to Read

This week, we'll be discussing: 

- Preface/Introduction
- Workflow basics
- Workflow scripts
- Workflow projects

### Physical Book

If you're reading the physical book, the chapters to read are: 

```{r echo=FALSE}
library(tidyverse)
read_csv("../physical-book-chapter-schedule.csv") %>% 
  filter(Meeting == 1) %>% 
  select(-Meeting) %>% 
  knitr::kable()
```

### Online Book

If you're reading the online book, the chapters to read are: 

```{r echo=FALSE}
library(tidyverse)
read_csv("../online-book-chapter-schedule.csv") %>% 
  filter(Meeting == 1) %>% 
  select(-Meeting) %>% 
  knitr::kable()
```

## Exercises 

All exercises refer to the online book chapters. The book exercises and online exercises may differ, so to make sure everyone is doing the same exercises, please refer to the online book chapters.

- Chapter 4: Section 4.4 #1-3
- Chapter 6: Section 6.3 #1-2

Exercises are reproduced below.

### Chapter 4 Exercises

1.  Why does this code not work?

    ```{r, error = TRUE}
    my_variable <- 10
    my_varıable
    ```

    Look carefully!
    (This may seem like an exercise in pointlessness, but training your brain to notice even the tiniest difference will pay off when programming.)
  
> The code doesn't work because the `i` in `my_variable` was replaced with a `ı`, a different character. The lesson here is that R takes everything you do very literally: `my_variable` is totally different from `my_varıable` or `My_variable` in R's eyes.    

2.  Tweak each of the following R commands so that they run correctly:

    ```{r, eval = FALSE}
    libary(tidyverse)
    ggplot(dota = mpg) + 
      geom_point(maping = aes(x = displ, y = hwy))
    
    fliter(mpg, cyl = 8)
    filter(diamond, carat > 3)
    ```
    
```{r, fig.align='center', fig.cap="A plot of displ by highway mpg"}
library(tidyverse)  # library not libary
ggplot(data = mpg) +  # data not dota
  geom_point(mapping = aes(x = displ, y = hwy))  # mapping not maping

filter(mpg, cyl == 8)  # filter not fliter, == not =
filter(diamonds, carat > 3) # diamonds, not diamond, is an available dataset 
```

> To find data sets in your environment: `data()` and hit Tab key inside of parens to look for all the data sets R has access to.

3.  Press Alt + Shift + K.
    What happens?

> Open up the keyboard shortcuts!    
    
How can you get to the same place using the menus?

> Go to Tools (or Help) -> Keyboard Shortcuts Help
    
### Chapter 6 Exercises

1.  Go to the RStudio Tips Twitter account, <https://twitter.com/rstudiotips> and find one tip that looks interesting. Practice using it!
    - *Maddie:* "The "reflow" command Ctrl+Shift+/ of @RStudio never ceases to amaze me. Sometimes I scramble the docs on purpose, just to watch it bounce back perfectly! Is there a way to reflow the Description field of DESCRIPTION similarly? #rstats"
    - *Jill:* "I learned  CTRL + SHIFT + M —> %>%"

    ```{r eval = FALSE, echo = FALSE}
    %>% 
    ```
    - *Simon:* Ctrl + Enter to run the line in editor, no selecting needed 
    - *Anna:* CTRL + U at the end of a line deletes all of the code from that line; Also, CTRL + ALT + i makes a new code chunk in an rmarkdown.
    - *Sam S:* [Visual markdown editor.](https://twitter.com/scarface_08944/status/1370352809275711490) (Rstudio version 1.4 or greater)
    - *Neelima:* Ctrl+1 2 3 etc moves cursor between panes in RStudio
    - *Paola:* Tweet from @ChelseaParlett. "In case you didn't know the shortcut: highlighting multiple lines in #rStudio and hitting tab to indent and Shift + tab to unindent MULTIPLE LINES AT ONCE I always do this when coding in python, but just never knew the command to do it in Rstudio!!"
    - *Sam T:* Alt(option) + Click & Drag highlights lines independently; Spellcheck : Edit -> Check spelling

2.  What other common mistakes will RStudio diagnostics report? Read <https://support.rstudio.com/hc/en-us/articles/205753617-Code-Diagnostics> to find out.
    - *Jill:*
        - Magic comments
        - Proper R style: space padding.

```{r, eval = FALSE}
ggplot(mpg) + 
  geom_point(aes(x = model, y = displ))

# not 

ggplot(mpg)+geom_point(aes(x=model,y=displ))
```
 

## Open Disussion

### What's the deal with the Rmd code chunk and its arguments?? 

#### Backticks, braces and how to use: 

- No brackets after a set of backticks gets you verbatim text:

```
verbatim code
```

- Define the engine in first argument after opening bracket, second argument is chunk name. Other arguments go after a comma after the name (see examples in source)

```{r pkgs, eval = FALSE, echo = FALSE}
library(tidyverse)
```

#### Code Chunk Arguments 

- See Yihui's post: https://yihui.org/knitr/options/ 
- Common ones (examples in source of this document): 
    * `eval`: **eval**uate/run the code or not? 
    * `echo`: display the code in the knitted document or not?
    * `fig.align`: how to **align** **fig**ures resulting from code? (left, right, center) 
    * `fig.cap`: Add a **cap**tion to a **fig**ure
    

#### How to use RStudio Cloud as a single user? 

- RStudio Cloud: A totally online version of RStudio, no downloading required. For more info see:
    - https://rstudio.cloud/personal/about
    - https://rstudio.cloud/learn/guide



